---
title: "IS 6489: Module 8 | BioImplants, Phase 2"
author: "Meag Tessmann"
date: "Nov 21 - Dec 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data and packages

```{r}
library(tidyverse)
library(caret)
library(scales) #moneys labels
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization


# Employee churn data
bi1 <- read_csv("bioimplants.csv")

# Employee productivity data
bi2 <- read_csv("additional_bioimplants.csv")

```

## Description of Data

- *age*:   age of employee in years.	
- *business_travel*: amount of job-related employee travel.            
- *department*:  department at the company in which the employee works.
- *distance_from_home*: travel distance to workplace from  employee's home, in miles.
- *education*:  employee's level of education.
- *education_field*:  disciplinary area of employee's education.
- *environment_satisfaction*: employee's satisfaction with the work environment.
- *gender*: Male/Female                     
- *job_involvement*:  employee's self-reported level of psychological and emotional  participation in work, profession, and company.
- *job_level*: the level of responsibility associated with the position, from entry level,  intermediate level, first-level management, middle-level management to top-level management (chiefs). 
- *job_role*: type of position at the company.
- *job_satisfaction*: self-reported job satisfaction.
- *marital_status*: Single, Married, Divorced.
- *monthly_income*: salary per month in $.          
- *num_companies_worked*: number of companies at which the employee has previously worked.
- *over_time*:  indicator variable for whether the employee has worked more than 40 hours per week.   
- *percent_salary_hike*: the percentage increase in  employee's salary in 2016 from 2015.     
- *performance_rating*:  supervisor's assessment of employee's performance.
- *relationship_satisfaction*:  employee's self-reported level of  satisfaction with relationships at work.
- *stock_option_level*:  indicator for the relative value of the stock options available to the employee.
- *total_working_years*: employee's total number of years in the workforce.
- *training_times_last_year*:  number of trainings attended by employee in 2016.
- *work_life_balance*: employee's self-reported work-life balance in current role at the company.     
- *years_at_company*:  number of years the employee has worked at the company.
- *years_in_current_role*: number of years the employee has worked at the company in current role.     
- *years_since_last_promotion*: number of years since the employee's last promotion.
- *years_with_curr_manager*:  number of years the employee has worked at the company with current supervisor.  
- *attrition*: indicator for whether the employee left the company in 2016.
- *employee_number*: employee ID number.
- *monthly rate*: employee monthly revenue production.

## Data Preparation

```{r}
# Join datasets

bi <- bi1 %>% 
  left_join(bi2, by = "employee_number")

# Use dummyVars() function along with predict() to make
# the numeric predictor matrix

num_frame <- dummyVars(attrition ~., 
                       fullRank=T, 
                       data = bi) %>% 
  predict(newdata = bi)

# Remove near zero variance predictors using nzv() function
num_frame <- num_frame[,-nzv(num_frame)]
 
# Impute with medians using preProcess() function together
# with predict()
num_frame <- preProcess(num_frame, "medianImpute") %>% 
  predict(num_frame)

# Remove the correlated variables over an arbitrary threshold.
num_frame <- num_frame[, -findCorrelation(cor(num_frame), cutoff = .75)]

# Make num_frame into a dat.frame
num_frame <- data.frame(num_frame)

# Look at variables and dimesnions of the dataset
glimpse(num_frame)

```


# Questions

## Q1: Estimate the current cost of attrition

```{r}
separation <- 500
acquisition <- 4900
placement <- 3500
workdays_per_year <- 240
workdays_per_month <- workdays_per_year / 12
workdays_position_open <- 40
workdays_during_onboarding <- 60
onboarding_efficiency <- .5 

direct_costs = separation+acquisition+placement

# Add attrition cost column to each employee row
bi <- bi %>% 
  mutate(
    attrition_cost = 
      (direct_costs - 
      (monthly_income / workdays_per_month*workdays_position_open) + 
      (bi$monthly_rate / workdays_per_month *
        (workdays_position_open+
        (workdays_during_onboarding * onboarding_efficiency)
        )
      ))
  )
   

# Calculate Average attrition - $45,636.18
mean(bi$attrition_cost, na.rm=TRUE)

# Check spread
bi %>% 
  filter(!is.na(attrition_cost)) %>% 
  summarize(
    q1 = quantile(attrition_cost, .25),
    q2 = quantile(attrition_cost, .5),
    mean = mean(attrition_cost),
    q3 = quantile(attrition_cost, .75),
    sd = sd(attrition_cost),
  )
  
  
  
```

> *Answer:* <br />
> Current average cost of attrition: $45,636.18

## Q2: Fit and use a model of churn

1. **Report average probability of churn for the baseline case.** 
2. **Report average probability of churn if there was no overtime.**

```{r}

# Logistic regression model  
set.seed(123)

# Remember: remove employee_number and monthly_rate.
num_frame_overtime <- num_frame %>% 
  select(-employee_number, -monthly_rate) 

# Train model
model_q2 <- train(
  x=num_frame_overtime,
  y=factor(bi$attrition),
  method = "glm",
  preProcess=c("center","scale")
)

# Probability of churn for overtime matrix: 0.1612245
pred_q2 <- as.data.frame(num_frame)
pred_q2$prediction_overtime = predict(model_q2, type = "prob", newdata = pred_q2)$Yes
mean(pred_q2$prediction_overtime, na.rm=TRUE)

# # Probability of churn for no overtime matrix: 0.103227
pred_q2$prediction_noOvertime = predict(model_q2, type = "prob", newdata = pred_q2 %>% mutate(over_timeYes=0))$Yes
mean(pred_q2$prediction_noOvertime, na.rm=TRUE)

# Check prediction var
summary(pred_q2$prediction_noOvertime)
summary(pred_q2$prediction_overtime)

```

> *Answer:* <br /><br />
> 1. Average probability of churn for the baseline case: 0.1612245<br/>
> 2. Average probability of churn if there was no overtime: 0.103227<br/>


## Q3: Calculate expected savings

```{r}

# To get you started:
ev <- data.frame(id = bi$employee_number,
                 attrition_cost = bi$attrition_cost,
                 cost_baseline = 0,
                 cost_new_policy =  .1 * bi$attrition_cost,
                 pLeave = pred_q2$prediction_overtime,
                 pLeave_noOvertime = pred_q2$prediction_noOvertime,
                 department = bi$department
                 )

# Add columns for expected cost for both overtime and no overtime policies, additionally expected savings column
ev <- ev %>% 
  mutate(
    expectedCost_overtime = pLeave*(attrition_cost+cost_baseline) + (1-pLeave)*cost_baseline,
    expectedCost_noOvertime = pLeave*(attrition_cost+cost_new_policy) + (1-pLeave)*cost_new_policy,
    expectedSavings = expectedCost_overtime - expectedCost_noOvertime
  )

# average cost of switching is $1,196.14
# total cost of switching is $1,672,197
ev %>% 
  filter(
    !is.na(expectedSavings)
  ) %>% 
  group_by(department) %>% 
  summarize(
    totalSavings = sum(expectedSavings),
    avgSavings = mean(expectedSavings),
    avgOvertime = mean(expectedCost_overtime),
    avgNoOvertime = mean(expectedCost_noOvertime),
    maxSavings = max(expectedSavings)
  )
head(ev)

```

> *Answer:*<br /><br />
> Total Cost with a no overtime policy: $1,672,197	<br />
> Average Cost per position with a no overtime policy: $1,196.14<br /><br />
> There seems to be an overall cost instead of savings for switching to a no overtime policy.

## Q4: Plot profit curve

**Comment on approximately many customers should be included in the no overtime policy.**

```{r}

ev_rank <- ev %>% 
  na.omit() %>% 
  arrange(desc(expectedSavings)) %>% 
  mutate(
    rank = dense_rank(desc(expectedSavings)),
    cumSum = cumsum(expectedSavings)
  ) 
ev_rank %>% 
  ggplot(aes(rank, cumSum)) + 
  geom_line() + 
  scale_y_continuous(labels=dollar)


  
```

> Answer: <br /><br/>
> About 50 employees have a positive expected savings. Of those 50, 8 currently are working overtime (calculated in question 5).

## Q5: Recommend

What should Angelica say in the phase 2 report? Please include details from the questions above (as well as from any additional analytics you might have done).


```{r}

# 5.3% of HR, 3.9% of R&D, 1.2% of Sales would generate savings based on new policy
ev <- ev %>% 
  mutate(
    changePolicy = ifelse(expectedSavings>0, 1, 0)
  ) 
ev %>% 
  filter(!is.na(expectedSavings)) %>% 
  group_by(department) %>% 
  summarize(
    mean(changePolicy)
  )


# Monthly rate and monthly income have high coefficients when predicting for a positive expected savings after a change in policy
colnames(ev)[1] <- "employee_number"
  
num_frame_q5 <- merge(ev, num_frame, by="employee_number")
num_frame_q5 <- num_frame_q5 %>% 
  select(-expectedCost_noOvertime, -expectedCost_overtime, -expectedSavings, -cost_new_policy, -cost_baseline, -attrition_cost) %>%
  mutate(
    changePolicy = factor(ifelse(is.na(changePolicy), 0,changePolicy))
  )

num_frame_q5 <- preProcess(num_frame_q5, "medianImpute") %>% 
  predict(num_frame_q5)
mean(is.na(num_frame_q5$changePolicy))

set.seed(123)
model_q5 <- train(
    changePolicy~., data=num_frame_q5,
    method = "glmnet",
    preProcess=c("center","scale"),
    na.action=na.exclude
)

# make a ordered graph, filtered to the highest coefficients
compare <- data.frame(variables = c("(Intercept)",model_q5$coefnames ))
compare$coefs <- c(as.numeric(as.character(coef(model_q5$finalModel, model_q5$bestTune$lambda))))
compare <- compare %>% filter(variables!="(Intercept)")
# Plot variables with highest coefficients
compare %>% 
  filter(abs(coefs)>.05) %>% 
  ggplot(aes(reorder(variables, coefs), coefs)) +
  geom_bar(stat = "identity",position="dodge") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# num_frame_kClust <- preProcess(num_frame_q5, c("center", "scale")) %>%  predict(num_frame_q5)

# Plot expected savings agaisnt highest coefficient variables
bi_plot <- merge(bi, ev %>% select(employee_number, expectedSavings), by="employee_number")

ggplot(bi_plot, aes(monthly_rate, expectedSavings)) + 
  geom_point()

ggplot(bi_plot, aes(monthly_income, expectedSavings)) + 
  geom_point()


  
# Partition data on monthly income > 10000 - there really isn't any other variables that could separate those in this subset.
  
num_frame_over10000 <- num_frame_q5 %>% 
  filter(monthly_income>10000, monthly_rate<10000)


set.seed(123)
model_over10000 <- train(
    changePolicy~., data=num_frame_over10000,
    method = "glmnet",
    preProcess=c("center","scale"),
    na.action=na.exclude
)

# make a ordered graph, filtered to the highest coefficients
compare_over10000 <- data.frame(variables = c("(Intercept)",model_over10000$coefnames ))
compare_over10000$coefs <- c(as.numeric(as.character(coef(model_over10000$finalModel, model_over10000$bestTune$lambda))))
compare_over10000 <- compare_over10000 %>% filter(variables!="(Intercept)")
# Plot variables with highest coefficients
compare_over10000 %>% 
  filter(abs(coefs)>.05) %>% 
  ggplot(aes(reorder(variables, coefs), coefs)) +
  geom_bar(stat = "identity",position="dodge") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Save money on 80% of employees by eliminating overtime - saving on average $245.8 per employee, 
bi_plot <- bi_plot %>% 
  mutate(
    over_time = factor(over_time)
  )
bi_plot %>% 
  filter(monthly_income>10000, monthly_rate<10000, education=="College", over_time=="Yes") %>% 
  summarize(
    mean_savings = mean(expectedSavings),
    count_savings = mean(expectedSavings>0),
  )





# check for those who already have overtime

num_frame_overtime <- num_frame_q5 %>% 
  filter(over_timeYes==1)


set.seed(123)
model_overtime <- train(
    changePolicy~., data=num_frame_overtime,
    method = "glmnet",
    preProcess=c("center","scale"),
    na.action=na.exclude
)

# make a ordered graph, filtered to the highest coefficients
compare_overtime <- data.frame(variables = c("(Intercept)",model_overtime$coefnames ))
compare_overtime$coefs <- c(as.numeric(as.character(coef(model_overtime$finalModel, model_overtime$bestTune$lambda))))
compare_overtime <- compare_overtime %>% filter(variables!="(Intercept)")
# Plot variables with highest coefficients
compare_overtime %>% 
  filter(abs(coefs)>.05) %>% 
  ggplot(aes(reorder(variables, coefs), coefs)) +
  geom_bar(stat = "identity",position="dodge") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

bi_plot %>% 
  filter(over_time=="Yes") %>% 
  ggplot(aes(monthly_rate, expectedSavings)) + 
  geom_point()

ggplot(bi_plot, aes(monthly_income, expectedSavings)) + 
  geom_point()


# Lose money 
bi_plot %>% 
  filter(monthly_income>10000, monthly_rate<10000, gender=="Male", over_time=="Yes") %>% 
  summarize(
    mean_savings = mean(expectedSavings),
    count_savings = mean(expectedSavings>0),
  )

bi_plot %>% 
  filter(over_time=="Yes", !is.na(over_time)) %>% 
  summarize(
    mean_change = mean(expectedSavings)
  )
summary(bi_plot$over_time)

# Find individual employees who would benefit from a change in policy
employee_list <- bi_plot %>% 
  filter(over_time=="Yes", expectedSavings>0)
employee_list %>% 
  summarize(
    totalSavings = sum(expectedSavings),
    count = NROW(expectedSavings)
  )
pred_q2 %>% 
  filter(employee_number %in% employee_list$employee_number) %>% 
  summarize(
    mean_predNo = mean(prediction_noOvertime),
    mean_predYes = mean(prediction_overtime)
  )


```



> Answer: <br /><br />
We looked at the cost associated with changing to a no overtime policy to help reduce employee churn. Taking into account direct costs (such as acquisition,) an employees monthly income, decreased onboarding efficiency, and average positions left open throughout the year, we calculated the average cost of losing an employee is $45,636.18. <br /><br />
While average probability of churn reduces from 16% to 10% if overtime is eliminated, we can conclude this policy change would cost the company \$1,672,197  when taking into account lost productivity and acquisition costs based on role and current monthly incomes. This would translate to \$1,196.14 per employee. <br /><br />
There are, however, a few employees that would benefit from a change in their individual overtime duties which would save the company an estimated $7,767.35 in preventing their departure. Their individual employee numbers are 124, 429, 531, 638, 661, 1005, 1167, 1608. Changing their overtime policy would result in a modest probability of churn reduction from 10% to 2%. 

---
- 
