---
title: "Case Study: Valmax Reality Case"
author: "Meag Tessmann"
date: "Oct 6-13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) 
library(janitor)
```

## Load data

```{r}
# Load data
set.seed(123)
valmax <- read_csv("valmax.csv")

# Inspect data
head(valmax)
glimpse(valmax)
summary(valmax)

```


```{r}
# no ticks here...
valmax <- clean_names(valmax, 'lower_camel')

# add difference column, clean data types 
valmax <- valmax %>% 
  mutate(
    difference = askingPrice - sellingPrice,
    broker = factor(broker),
    house = as.integer(house)
  )

```


## Questions

### Q1

Plot the difference between asking vs selling price by broker. Does there appear to be a difference?

<p style="color:blue">
Valmax does appear to have a smaller difference than other brokers, with a mean $4653 lower than other brokers and median $6500 lower. 
</p>
```{r}

# get an idea of the data
valmax %>%
  group_by(broker) %>%
  ggplot(aes(broker, difference)) + 
  geom_boxplot()

sum_stats <- valmax %>%
  group_by(broker) %>%
  summarize(
    mean(difference),
    median(difference), 
    sd(difference),
    NROW(broker)
  )

sum_stats
```

### Q2

Calculate and interpret 95% CIs for mean difference by broker. 

<p style="color: blue"> 
While there is a lot of overlap, Valmax does have a lower confidence interval both on the upper and lower ends, suggesting that they do have a slightly lower difference between asking and selling price than other brokers.
</p>

```{r}
# seperate out summary stats to one object per broker
v_stats <- sum_stats %>%
  filter(broker=='valmax')

o_stats <- sum_stats %>%
  filter(broker=='others')
 
# create function to compute CIs
calc_ci_upper <- function(s_mean, s_sd) {
  s_mean+1.96*(s_sd)
}
calc_ci_lower <- function(s_mean, s_sd) {
  s_mean-1.96*(s_sd)
}

# computer CI upper and lower  bounds for each broker
v_ci_upper <- calc_ci_upper(v_stats$`mean(difference)`, v_stats$`sd(difference)`)
o_ci_upper <- calc_ci_upper(o_stats$`mean(difference)`, o_stats$`sd(difference)`)
v_ci_lower <- calc_ci_lower(v_stats$`mean(difference)`, v_stats$`sd(difference)`)
o_ci_lower <- calc_ci_lower(o_stats$`mean(difference)`, o_stats$`sd(difference)`)

v_ci_lower
v_ci_upper
o_ci_lower
o_ci_upper

# plot confidence intervals
valmax %>% 
  ggplot(aes(difference, color=broker)) + 
  geom_density() + 
  geom_vline(xintercept = v_ci_lower, lty=2, color='light sea green') + 
  geom_vline(xintercept = v_ci_upper, lty=2, color='light sea green') + 
  geom_vline(xintercept = o_ci_lower, lty=2, color='coral') + 
  geom_vline(xintercept = o_ci_upper, lty=2, color='coral') + 
  theme_minimal()
  


```

### Q3

Compute a t-statistic for mean difference by broker. 

```{r}

# break apart observations by broker

val_obs <- valmax %>%
  filter(broker=="valmax") %>%
  select(difference)
val_obs <- as.vector(val_obs$difference)
oth_obs <- valmax %>%
  filter(broker=='others') %>%
  select(difference)
oth_obs <- as.vector(oth_obs$difference)

# calculate t-statistic, assuming 'others' broker as the null distribution

t_stat <- t.test(val_obs, oth_obs)

t_stat$statistic

```

### Q4

What is the p-value associated with the t-statistic you have computed? Use a 2-tailed test by doubling the p-value obtained with the `pt()` function. (Use the built in t.test function to check your work. Your results should be very close.)

```{r}

#degrees of freedom 
df <- v_stats$`NROW(broker)`-1

#calculate 2-tailed p-value for t-statistic
pt(t_stat$statistic, df)*2

#compare to built in t-test function's p-value calculation
t_stat$p.value

```

### Q5

Are the clients' complaints justified? Write a paragraph in which you weigh the evidence and argue a position. 

<p style="color: blue">
We have sufficient evidence (α=.05) to reject the null hypothesis: 'The mean difference between asking and selling prices when Valmax is the broker is equal to the mean of other brokers' differences'. There is only a 4.3% probability that there is no difference between Valmax and others' difference in asking and selling prices. This implies Valmax clients are justified in either or both of two complaints: being advised to set an asking price for their house too low or to buy a house for too much money. 
</p>



